### 2017-01-21 second-order-spsa-3.tex
chapter 4 conditions changed

 \item \textbf{Execution}: \\
 \For{$n \gets 0,1,2, \ldots, $} {
 \begin{itemize}
 \item Generate $\{\Delta_n^{i}, i=1,\ldots,N\}$,$\{\widehat\Delta_n^{i}, i=1,\ldots,N\}$ independent of $\{\Delta_m,\widehat\Delta_m, m=0,1,\ldots,n-1\}$.  
 \item For any $i=1,\ldots,N$, $\Delta_n^{i}$,$\widehat\Delta_n^{i}$  satisfy condition (C17) of chapter \ref{sec:2spsa-results},(most popular distribution being used is symmetric Bernoulli distribution, $\pm 1$ w.p. $1/2$).
\begin{itemize}
\item \textbf{\emph {Function evaluation 1} }\\ \hspace{4em} Obtain $y_n^+ = f(\theta_n+\delta_n \Delta_n +\delta_n \widehat\Delta_n) + \xi_n^+$.
\item \textbf{\emph {Function evaluation 2} }\\ \hspace{4em} Obtain $y_n^- = f(\theta_n -\delta_n\Delta_n-\delta_n\widehat\Delta_n) + \xi_n^-$.
\item \textbf{\emph {Function evaluation 3} }\\ \hspace{4em} Obtain $y_n = f(\theta_n) + \xi_n$.
\item \textbf{\emph {Newton step}}\\ \hspace{4.2em} Update the parameter and Hessian as follows:
%\begin{align*}
%		\theta_{n+1} = & \; \theta_n - a_n \Upsilon(\overline H_n)^{-1}\widehat\nabla f(\theta_n), \\
%\overline H_n = &\; (1-b_{n})  \overline H_{n-1} + b_{n} ( \widehat H_n - \widehat \Psi_n),
%\end{align*}
\begin{align} \label{eq:2spsa-3-theta}
\theta_{n+1} = \,& \theta_n - a_n \Upsilon(\overline H_n)^{-1}\widehat\nabla f(\theta_n),\\
\overline H_n =\,& (1-b_{n})  \overline H_{n-1} + b_{n} ( \widehat H_n - \widehat \Psi_n),\label{eq:2spsa-3-ih}
\end{align}
where $\widehat H_n$ and $\widehat \Psi_n$ are chosen according to \eqref{eq:2spsa-ber} and \eqref{eq:psinhatspsa3}, respectively. 
\end{itemize}

@@

Further, $\delta_n, n\geq 0$ is a sequence of diminishing positive real numbers and 
$\Delta_n = (\Delta_n^1,\ldots,\Delta_n^N)\tr$ , $\widehat\Delta_n = (\widehat\Delta_n^1,\ldots,\widehat\Delta_n^N)\tr$ are the perturbation random vectors at instant $n$ with $\Delta_n^i,i=1,\ldots,N$, and $\widehat\Delta_n^i,i=1,\ldots,N$ being independent identically distributed (i.i.d), mean-zero random variables having finite inverse moments of order greater than 2 and satisfying the condition (C17) of section \ref{sec:2spsa-results}.

@@

The optimal choice for $b_n$ in \eqref{eq:2spsa-3-ih}  is the following:
\begin{align}
\label{eq:spsawieghts}
b_i  = \delta_i^{4}/\sum\limits_{j=0}^{i} \delta_j^{4}.
\end{align}
The main idea behind the above choice is provided below.
From \eqref{eq:spsahnhat}, we can infer that
\begin{align*}
\E \parallel \widehat H_n \parallel^2 \le \dfrac{C}{\delta_n^4} \text{ for some } C<\infty. 
\end{align*} 
This is because the third term in \eqref{eq:spsahnhat} vanishes asymptotically, while the fourth term there dominates asymptotically. Moreover, the noise factors in the fourth term in \eqref{eq:spsahnhat} are bounded above due to (C22)  and independent of $n$, leaving the $\delta_n^2$ term in the denominator there.

@@

\item For any $\tau >0$ and nonempty $S \subseteq \{1,\ldots,N\}$, there exists a $\rho'(\tau,S)>\tau$ such that 
$$ \limsup_{n\rightarrow \infty} \left| \dfrac{\sum_{i \notin S} (\theta-\theta^*)_i \bar f_{ni}(\theta)}{\sum_{i \in S} (\theta-\theta^*)_i \bar f_{ni}(\theta)}               \right| < 1 \text{ a.s.}$$
for all $|(\theta-\theta^*)_i| < \tau$ when $i \notin S$ and $|(\theta-\theta^*)_i| \ge \rho'(\tau,S)$ when $i\in S$.
\item For some $\alpha_0, \alpha_1, \alpha_2 >0$ and for all $n,l,m$, $\E {\xi_n}^{2} \le \alpha_0$, $\E {\xi_n^{\pm}}^{2} \le \alpha_0$, $\E f(\theta_n)^{2} \le \alpha_1$,  $\E f(\theta_n + \delta_n \Delta_n + \delta_n \widehat \Delta_n)^{2},\E f(\theta_n - \delta_n \Delta_n - \delta_n \widehat \Delta_n)^{2} \le \alpha_1$, 
%$\E \left[ f(\theta_n + \delta_n \Delta_n + \delta_n \widehat \Delta_n)^{2}/ (\Delta_{nl}\widehat\Delta_{nm})^2 \mid \F_n\right]$,
$\E \left[| f(\theta_n + \delta_n \Delta_n + \delta_n \widehat \Delta_n)/ (\Delta_{nl}\widehat\Delta_{nm})|^{2+\alpha_2} \mid \F_n\right]$,
%$\E \left[ f(\theta_n - \delta_n \Delta_n - \delta_n \widehat \Delta_n)^{2}/ (\Delta_{nl}\widehat\Delta_{nm})^2 \mid \F_n\right]$,
$\E \left[ |f(\theta_n - \delta_n \Delta_n - \delta_n \widehat \Delta_n)/ (\Delta_{nl}\widehat\Delta_{nm})|^{2+\alpha_2} \mid \F_n\right]$,
 $\E\left[\left. (\xi_n^+ + \xi_n^- - 2 \xi_n)^2/(\Delta_{nl}\widehat\Delta_{nm})^2 \right| \F_n \right] \le \alpha_1$ 
and $\E \left(\left\| \Upsilon(\overline H_n) \right\|^2 \mid \F_n\right) \le \alpha_1$. 
\item  $\delta_n = \frac{\delta_0}{(n+1)^{\varsigma}}$, where $\delta_0 > 0$ and $0 < \varsigma \le 1/8$.
\end{pvn}
The reader is referred to Section II-B of \cite{prashanth2015rdsa} for a detailed discussion of the above assumptions. We remark here that (C14)-(C21) are identical to those in \cite{prashanth2015rdsa}, while (C22) and (C23) introduce minor additional requirements on $\left\| \Upsilon(\overline H_n) \right\|^2$ and $\delta_n$, respectively and these are inspired from \cite{spall-jacobian}.

\begin{lemma}(\textbf{2SPSA-3-IH Bias in Hessian estimate})
\label{lemma:2spsa-bias}
Under (C14)-(C23), with $\widehat H_n$ defined according to  \eqref{eq:2spsa-ber} , we have a.s. that\footnote{Here $\widehat H_n(i,j)$ and $\nabla^2_{i,j}f(\cdot)$ denote the $(i,j)$th entry in the Hessian estimate $\widehat H_n$ and the true Hessian $\nabla^2 f(\cdot)$, respectively.}, for $i,j = 1,\ldots,N$,
\begin{align}
\left|\E\left[
\left. \widehat H_n(i,j) \right| \F_n \right] - \nabla^2_{ij} f(\theta_n)\right| = O(\delta_n^2).
\end{align} 
\end{lemma}
\begin{proof}
See Proposition 4.2 in \cite{bhatnagar2015simultaneous}.
\end{proof}

\begin{theorem}(\textbf{2SPSA-3-IH Strong Convergence of Hessian})
\label{thm:2spsa-H}
Under (C14)-(C23), we have that 
$$\theta_n \rightarrow \theta^*, \overline H_n \rightarrow \nabla^2 f(\theta^*) \text{ a.s. as } n\rightarrow \infty.$$ 
In the above, $\theta_n$ and $\overline H_n$ are updated according to \eqref{eq:2spsa-3-theta} and \eqref{eq:2spsa-3-ih}, respectively, $\widehat H_n$ defined according to \eqref{eq:2spsa-ber} and the step-sizes $b_n$ are chosen as suggested in \eqref{eq:spsawieghts}. 
\end{theorem}
\begin{proof}
The proof is similar to the proof of the Theorem \ref{thm:2rdsa-H} and is therefore omitted.
\end{proof}

@@

\item The operator $\Upsilon$ is chosen such that $\E(\parallel \Upsilon(\overline H_n) - \overline H_n\parallel^2) = o(e^{-2wn^{1-r}/(1-r)})$ and $\parallel \Upsilon(H) - H \parallel^2 / (1+\parallel H \parallel^2)$ is uniformly bounded.

@@

Assume (C17), (C23), (C24) and (C25) and also that the setting is noise-free. 

@@

@@

Now consider the second case where $lim sup_{n \to \infty} \frac{d(n)}{n^{2r}} = \infty$. There exists a subsequence $\{i_1,\ldots,i_n\}$ such that $ \frac{d(i_n)}{i_n^{2r}} \ge \frac{d(i)}{i^{2r}}$ as $i \le i_n$. Then for some $C >0$ the RHS of \eqref{eq:thmwts} satisfies the following 
\begin{align}
\text{RHS}(i_n)  \text{of}  \eqref{eq:thmwts} &= O(1) e^{-2 \Gamma_{1i_n}} +  O(1) e^{-2 b_0 i_n^{1-r}/(1-r)} \left[\sum_{i=1}^{i_n} e^{2 b_0 i^{1-r}/(1-r)} \frac{b_0^2}{i^{2r}} \E (\Lambda_{n} \otimes \Lambda_{n}) \right]\nonumber \\
&\le C e^{-2b_0n^{1-r} / {1-r}} \left[ 1 + O(1) \sum_{i =1}^{i_n} \frac{d(i)}{i^{2r}}\right]\nonumber \\
&\le C e^{-2b_0n^{1-r} / {1-r}} \left[ 1 + \frac{d(i_n)}{{i_n}^{2r}}\right].
\label{eq:spsaquad-bigocont5}
\end{align}
From the equation \eqref{eq:spsaquad-bigocont}, the  LHS of \eqref{eq:thmwts} is $d(i_n) e^{-2b_0 i_n^{1-r} / {1-r}}$. By substituting \eqref{eq:spsaquad-bigocont} and \eqref{eq:spsaquad-bigocont5}, in $\frac{RHS{i_n}}{LHS{i_n}}$ then $\frac{RHS{i_n}}{LHS{i_n}} \to 0$ as $n \to \infty$. It is contradiction since $LHS(n) = RHS(n)$ for all $n$. Hence the claim $\text{trace}[\E (\Lambda_n \tr \Lambda_n)] = O(e^{-2b_0n^{1-r} / {1-r}})$ follows.
\end{proof}

>>>>

 \item \textbf{Execution}: \\
 \For{$n \gets 0,1,2, \ldots, $} {
 \begin{itemize}
 \item Generate $\{\Delta_n^{i}, i=1,\ldots,N\}$,$\{\widehat\Delta_n^{i}, i=1,\ldots,N\}$ independent of $\{\Delta_m,\widehat\Delta_m, m=0,1,\ldots,n-1\}$.  
 \item For any $i=1,\ldots,N$, $\Delta_n^{i}$,$\widehat\Delta_n^{i}$  satisfy condition (A23) of chapter \ref{sec:2spsa-results},(most popular distribution being used is symmetric Bernoulli distribution, $\pm 1$ w.p. $1/2$).
\begin{itemize}
\item \textbf{\emph {Function evaluation 1} }\\ \hspace{4em} Obtain $y_n^+ = f(\theta_n+\delta_n \Delta_n +\delta_n \widehat\Delta_n) + \xi_n^+$.
\item \textbf{\emph {Function evaluation 2} }\\ \hspace{4em} Obtain $y_n^- = f(\theta_n -\delta_n\Delta_n-\delta_n\widehat\Delta_n) + \xi_n^-$.
\item \textbf{\emph {Function evaluation 3} }\\ \hspace{4em} Obtain $y_n = f(\theta_n) + \xi_n$.
\item \textbf{\emph {Newton step}}\\ \hspace{4.2em} Update the parameter and Hessian as follows:
%\begin{align*}
%		\theta_{n+1} = & \; \theta_n - a_n \Upsilon(\overline H_n)^{-1}\widehat\nabla f(\theta_n), \\
%\overline H_n = &\; (1-b_{n})  \overline H_{n-1} + b_{n} ( \widehat H_n - \widehat \Psi_n),
%\end{align*}
\begin{align} \label{eq:2spsa-3-theta}
\theta_{n+1} = \,& \theta_n - a_n \Upsilon(\overline H_n)^{-1}\widehat\nabla f(\theta_n),\\
\overline H_n =\,& (1-b_{n})  \overline H_{n-1} + b_{n} ( \widehat H_n - \widehat \Psi_n),\label{eq:2spsa-3-ih}
\end{align}
where $\widehat H_n$ and $\widehat \Psi_n$ are chosen according to \eqref{eq:2spsa-ber} and \eqref{eq:psinhatspsa3}, respectively. 
\end{itemize}

@@

Further, $\delta_n, n\geq 0$ is a sequence of diminishing positive real numbers and 
$\Delta_n = (\Delta_n^1,\ldots,\Delta_n^N)\tr$, $\widehat\Delta_n = (\widehat\Delta_n^1,\ldots,\widehat\Delta_n^N)\tr$ are the perturbation random vectors at instant $n$ with $\Delta_n^i,i=1,\ldots,N$, and $\widehat\Delta_n^i,i=1,\ldots,N$ being independent identically distributed (i.i.d), mean-zero random variables having finite inverse moments of order greater than 2 and satisfying the condition (A23) of section \ref{sec:2spsa-results}.

@@

The optimal choice for $b_n$ in \eqref{eq:2spsa-3-ih}  is the following:
\begin{align}
\label{eq:spsawieghts}
b_i  = \delta_i^{4}/\sum\limits_{j=0}^{i} \delta_j^{4}.
\end{align}
The main idea behind the above choice is provided below.
From \eqref{eq:spsahnhat}, we can infer that
\begin{align*}
\E \parallel \widehat H_n \parallel^2 \le \dfrac{C}{\delta_n^4} \text{ for some } C<\infty. 
\end{align*} 
This is because the third term in \eqref{eq:spsahnhat} vanishes asymptotically, while the fourth term there dominates asymptotically. Moreover, the noise factors in the fourth term in \eqref{eq:spsahnhat} are bounded above due to (A28)  and independent of $n$, leaving the $\delta_n^2$ term in the denominator there.

@@

\item For any $\tau >0$ and nonempty $S \subseteq \{1,\ldots,N\}$, there exists a $\rho'(\tau,S)>\tau$ such that 
$$ \limsup_{n\rightarrow \infty} \left| \dfrac{\sum_{i \notin S} (\theta-\theta^*)_i \bar f_{ni}(\theta)}{\sum_{i \in S} (\theta-\theta^*)_i \bar f_{ni}(\theta)}               \right| < 1 \text{ a.s.}$$
for all $|(\theta-\theta^*)_i| < \tau$ when $i \notin S$ and $|(\theta-\theta^*)_i| \ge \rho'(\tau,S)$ when $i\in S$.
\item For some $\alpha_0, \alpha_1, \alpha_2 >0$ and for all $n,l,m$, $\E {\xi_n}^{2} \le \alpha_0$, $\E {\xi_n^{\pm}}^{2} \le \alpha_0$, $\E f(\theta_n)^{2} \le \alpha_1$,  $\E f(\theta_n + \delta_n \Delta_n + \delta_n \widehat \Delta_n)^{2},\E f(\theta_n - \delta_n \Delta_n - \delta_n \widehat \Delta_n)^{2} \le \alpha_1$, 
%$\E \left[ f(\theta_n + \delta_n \Delta_n + \delta_n \widehat \Delta_n)^{2}/ (\Delta_{nl}\widehat\Delta_{nm})^2 \mid \F_n\right]$,
$\E \left[| f(\theta_n + \delta_n \Delta_n + \delta_n \widehat \Delta_n)/ (\Delta_{nl}\widehat\Delta_{nm})|^{2+\alpha_2} \mid \F_n\right]$,
%$\E \left[ f(\theta_n - \delta_n \Delta_n - \delta_n \widehat \Delta_n)^{2}/ (\Delta_{nl}\widehat\Delta_{nm})^2 \mid \F_n\right]$,
$\E \left[ |f(\theta_n - \delta_n \Delta_n - \delta_n \widehat \Delta_n)/ (\Delta_{nl}\widehat\Delta_{nm})|^{2+\alpha_2} \mid \F_n\right]$,
 $\E\left[\left. (\xi_n^+ + \xi_n^- - 2 \xi_n)^2/(\Delta_{nl}\widehat\Delta_{nm})^2 \right| \F_n \right] \le \alpha_1$ 
and $\E \left(\left\| \Upsilon(\overline H_n) \right\|^2 \mid \F_n\right) \le \alpha_1$. 
\item  $\delta_n = \frac{\delta_0}{(n+1)^{\varsigma}}$, where $\delta_0 > 0$ and $0 < \varsigma \le 1/8$.
\end{pvn}
The reader is referred to Section II-B of \cite{prashanth2015rdsa} for a detailed discussion of the above assumptions. We remark here that (A20)-(A27) are identical to those in \cite{prashanth2015rdsa}, while (A28) and (A29) introduce minor additional requirements on $\left\| \Upsilon(\overline H_n) \right\|^2$ and $\delta_n$, respectively and these are inspired from \cite{spall-jacobian}.

\begin{lemma}(\textbf{2SPSA-3-IH Bias in Hessian estimate})
\label{lemma:2spsa-bias}
Under (A20)-(A29), with $\widehat H_n$ defined according to  \eqref{eq:2spsa-ber} , we have a.s. that\footnote{Here $\widehat H_n(i,j)$ and $\nabla^2_{i,j}f(\cdot)$ denote the $(i,j)$th entry in the Hessian estimate $\widehat H_n$ and the true Hessian $\nabla^2 f(\cdot)$, respectively.}, for $i,j = 1,\ldots,N$,
\begin{align}
\left|\E\left[
\left. \widehat H_n(i,j) \right| \F_n \right] - \nabla^2_{ij} f(\theta_n)\right| = O(\delta_n^2).
\end{align} 
\end{lemma}
\begin{proof}
See Proposition 4.2 in \cite{bhatnagar2015simultaneous}.
\end{proof}

\begin{theorem}(\textbf{2SPSA-3-IH Strong Convergence of Hessian})
\label{thm:2spsa-H}
Under (A20)-(A29), we have that 
$$\theta_n \rightarrow \theta^*, \overline H_n \rightarrow \nabla^2 f(\theta^*) \text{ a.s. as } n\rightarrow \infty.$$ 
In the above, $\theta_n$ and $\overline H_n$ are updated according to \eqref{eq:2spsa-3-theta} and \eqref{eq:2spsa-3-ih}, respectively, $\widehat H_n$ defined according to \eqref{eq:2spsa-ber} and the step-sizes $b_n$ are chosen as suggested in \eqref{eq:spsawieghts}. 
\end{theorem}
\begin{proof}
The proof is similar to the proof of the Theorem \ref{thm:2rdsa-H} and is therefore omitted.
\end{proof}

@@

\item The operator $\Upsilon$ is chosen such that $\E(\parallel \Upsilon(\overline H_n) - \overline H_n\parallel^2) = o(e^{-2wn^{1-r}/(1-r)})$ and \\$\parallel \Upsilon(H) - H \parallel^2 / (1+\parallel H \parallel^2)$ is uniformly bounded.

@@

Assume (A23), (A29), (A30) and (A31) and also that the setting is noise-free. 

@@

By rewriting the \eqref{eq:2spsa-3-ih} in the following form
\begin{align}\label{eq:hhatexpanded}
\overline H_{n} &= \overline H_{n-1} - b_{n} ( \overline H_{n-1} - \widehat H_n + \widehat \Psi_n)\nonumber\\
&= \overline H_{n-1} - b_{n} ( \overline H_{n-1} - H^* + \widehat \Psi_n - \Psi_n (H^*))
The above equation can be thought as a stochastic approximation analogue of \eqref{eq:2spsa-3-ih}.

@@

Now consider the second case where $lim sup_{n \to \infty} \frac{d(n)}{n^{2r}} = \infty$. There exists a subsequence $\{i_1,\ldots,i_n\}$ such that $ \frac{d(i_n)}{i_n^{2r}} \ge \frac{d(i)}{i^{2r}}$ as $i \le i_n$. Then for some $C >0$ the RHS of \eqref{eq:thmwts} satisfies the following 
\begin{align}
\text{RHS}(i_n)  \text{of}  \eqref{eq:thmwts} &= O(1) e^{-2 \Gamma_{1i_n}} +  O(1) e^{-2 b_0 i_n^{1-r}/(1-r)} \left[\sum_{i=1}^{i_n} e^{2 b_0 i^{1-r}/(1-r)} \frac{b_0^2}{i^{2r}} \E (\Lambda_{n} \otimes \Lambda_{n}) \right]\nonumber \\
&\le C e^{-2b_0n^{1-r} / {1-r}} \left[ 1 + O(1) \sum_{i =1}^{i_n} \frac{d(i)}{i^{2r}}\right]\nonumber \\
&\le C e^{-2b_0n^{1-r} / {1-r}} \left[ 1 + \frac{d(i_n)}{{i_n}^{2r}}\right].
\label{eq:spsaquad-bigocont5}
\end{align}
From the equation \eqref{eq:spsaquad-bigocont}, the  LHS of \eqref{eq:thmwts} is $d(i_n) e^{-2b_0 i_n^{1-r} / {1-r}}$. By substituting \eqref{eq:spsaquad-bigocont} and \eqref{eq:spsaquad-bigocont5}, in $\frac{RHS(i_n)}{LHS(i_n)}$ then $\frac{RHS(i_n)}{LHS(i_n)} \to 0$ as $n \to \infty$. It is contradiction since $LHS(n) = RHS(n)$ for all $n$. Hence the claim $\text{trace}[\E (\Lambda_n \tr \Lambda_n)] = O(e^{-2b_0n^{1-r} / {1-r}})$ follows.
\end{proof}
